Convective heat transport in stratified atmospheres at low and high Mach number

### 2016-11-17 tex/nusselt_paper.tex
as-->because, grad-->partial_z, extra comma

In RB convection, there exist two primary control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity).  These numbers coupled with the aspect ratio of 
the physical domain and the boundary conditions
determine the dynamics of the convection.  In stratified atmospheres, 
in addition to specifying the equation of state and
fundamental properties of the gas, the two control parameters of 
RB convection are joined by the degree of
stratification across the domain and the characteristic 
Ma of the convective flows.  
Polytropically stratified atmospheres, such as those used in 
early studies \cite{graham1975, chan&all1982, hurlburt&all1984, 
cattaneo&all1990, cattaneo&all1991, brummell&all1996}, are an ideal extension of
RB convection into the stratified realm as the two additional 
control parameters are directly linked to
basic properties of the atmosphere.  The density stratification is 
set by the number of density scale heights (\nrho)
the atmosphere spans, and Ma is controlled 
by the superadiabatic excess ($\epsilon$),
the deviation of the polytropic index from the adiabatic polytropic 
index \cite{graham1975}.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We exmine the simplest stratified extension of RB by studying a
fluid composed of monatomic ideal gas particles with an adiabatic 
index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
This is consistent with the approach used in earlier work.
The initial atmosphere is a plane-parallel polytrope in which 
the gravitational acceleration and conductive flux, 
$\bm{F}_{\text{cond,0}} = -\kappa \grad T_0$, do not vary with depth. To
achieve the latter condition, both $\kappa$ and $\grad T_0$ are constant.
Under these assumptions, satisfying hydrostatic 
equilibrium produces a stratification of
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
T_0(z)    &= T_{t}(z_0 - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$ and
the superadiabatic excess is $\epsilon$ which sets the scale 
of the entropy gradient ($\grad S_0 \propto -\epsilon$).
A significant advance of this work is the ability to study large 
and small values of $\epsilon$, as will be discussed.
Thermodynamic variables are nondimensionalized at the 
top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
By this choice, the non-dimensional length scale is the inverse 
temperature gradient scale and the timescale is the isothermal 
sound crossing time of this unit length.
The height $z$ increases upwards within $[0, L_{z}]$, 
where $L_{z} = e^{n_{\rho}/m} - 1$ is
determined by $n_\rho$ and $\epsilon$.
The characteristic timescale of convective dynamics
is related to the atmospheric buoyancy time, 
$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
Throughout this letter, we use buoyancy time units and 
choose $n_{\rho} = 3$ such that the 
initial density varies by a factor of 20.
All atmospheres studied here have an aspect ratio of 4, 
such that $L_x = 4L_z$.

>>>>

In RB convection, there exist two primary control parameters: 
the Rayleigh number (Ra, the ratio of
buoyant driving to diffusive damping) and the Prandtl number 
(Pr, the ratio of viscous to thermal
diffusivity).  These numbers coupled with the aspect ratio of 
the physical domain and the boundary conditions
determine the dynamics of the convection.  In stratified atmospheres, 
in addition to specifying the equation of state and
fundamental properties of the gas, the two control parameters of 
RB convection are joined by the degree of
stratification across the domain and the characteristic 
Ma of the convective flows.  
Polytropically stratified atmospheres, such as those used in 
early studies \cite{graham1975, chan&all1982, hurlburt&all1984, 
cattaneo&all1990, cattaneo&all1991, brummell&all1996}, are an ideal extension of
RB convection into the stratified realm because the two additional 
control parameters are directly linked to
basic properties of the atmosphere.  The density stratification is 
set by the number of density scale heights (\nrho)
the atmosphere spans, and Ma is controlled 
by the superadiabatic excess ($\epsilon$),
the deviation of the polytropic index from the adiabatic polytropic 
index \cite{graham1975}.

@@

\section{Experiment} 
\refstepcounter{section}
\label{sec:experiment}
We exmine the simplest stratified extension of RB by studying a
fluid composed of monatomic ideal gas particles with an adiabatic 
index of $\gamma = 5/3$ and whose equation of state is $P = R\rho T$. 
This is consistent with the approach used in earlier work.
The initial atmosphere is a plane-parallel polytrope in which 
the gravitational acceleration and conductive flux, 
$\bm{F}_{\text{cond,0}} = -\kappa \partial_z T_0$, do not vary with depth. To
achieve the latter condition, both $\kappa$ and $\partial_z T_0$ are constant.
Under these assumptions, satisfying hydrostatic 
equilibrium produces a stratification of
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(z_0 - z)^m, \\
T_0(z)    &= T_{t}(z_0 - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m = m_{\text{ad}} - \epsilon$ is the polytropic index.
The adiabatic polytropic index is $m_{\text{ad}} \equiv (\gamma-1)^{-1}$, and
the superadiabatic excess is $\epsilon$ which sets the scale 
of the entropy gradient ($\partial_z S_0 \propto -\epsilon$).
A significant advance of this work is the ability to study large 
and small values of $\epsilon$, as will be discussed.
Thermodynamic variables are nondimensionalized at the 
top of the atmosphere as  $P_0(L_z) = \rho_0(L_z) = T_0(L_z) = 1$, 
requiring $z_0 \equiv L_z + 1$ and $R = T_{t} = \rho_{t} = 1$.
By this choice, the non-dimensional length scale is the inverse 
temperature gradient scale and the timescale is the isothermal 
sound crossing time of this unit length.
The height $z$ increases upwards within $[0, L_{z}]$, 
where $L_{z} = e^{n_{\rho}/m} - 1$ is
determined by $n_\rho$ and $\epsilon$.
The characteristic timescale of convective dynamics
is related to the atmospheric buoyancy time, 
$t_{\text{b}} = \sqrt{L_z/g\epsilon}$, with $g = (m+1)$.
Throughout this letter, we use buoyancy time units and 
choose $n_{\rho} = 3$ such that the 
initial density varies by a factor of 20.
All atmospheres studied here have an aspect ratio of 4, 
such that $L_x = 4L_z$.
